Computable Poincar\'e--Friedrichs constants for the $L^{p}$ de Rham complex over convex domains and domains with shellable triangulations

### 2024-10-06 main.tex
more corrections

%\subsection{\cye{Potential for the gradient operator and the corresponding Poincar\'e--Friedrichs inequality}} \label{section:intro_grad}

@@

\cye{Let a vector field $\bff$ that is a distributional curl of an object from $\bfW^{p}(\curl,\Omega)$, $\bff \in \curl \bfW^{p}(\curl,\Omega)$, and scalar-valued field $f$ that is a distributional divergence of an object from $\bfW^{p}(\divergence,\Omega)$, $f \in \divergence \bfW^{p}(\divergence,\Omega)$, be given.} 
We are interested in constants for inequalities such as 
\begin{subequations}\label{math:pot_PF_curl_div} \begin{align}
    \min\limits_{ \substack{ \bfu \in \bfW^{p}(\curl,\Omega) \\ \curl \bfu = \bff } } 
    \| \bfu \|_{L^{p}(\Omega)}
    &\leq 
    C_{\curl,\Omega,p}
    \| \bff \|_{L^{p}(\Omega)} \label{math:pot_PF_curl}
    ,
    \\
    \min\limits_{ \substack{ \bfu \in \bfW^{p}(\divergence,\Omega) \\ \divergence \bfu = f } } 
    \| \bfu \|_{L^{p}(\Omega)}
    &\leq 
    C_{\divergence,\Omega,p}
    \| f \|_{L^{p}(\Omega)} \label{math:pot_PF_div}
    ,
\end{align}\end{subequations}
and how these constants can be estimated by the operator norms of {\em linear} potential operators 
\begin{align*}
    &\Phi_{\curl} : \curl \bfW^{p}(\curl,\Omega) \rightarrow \bfW^{p}(\curl,\Omega)
    ,
    \\
    &\Phi_{\divergence} : \divergence \bfW^{p}(\divergence,\Omega) \rightarrow \bfW^{p}(\divergence,\Omega)
    .
\end{align*}
The fundamental difference \cye{with respect to Section~\ref{section:intro_grad}} is that the curl and divergence have infinite-dimensional kernels. In Section~\ref{section:intro_grad}, the kernel of the gradient is the one-dimensional space of constant functions. It is thus trivially complemented for all $p$, with a canonical choice of projection. By contrast, in the Banach space case $p \neq 2$, it is not immediately evident that the kernels of the curl and divergence operators are complemented, and a canonical projection only exists here in the Hilbert setting $p=2$. \cye{Actually,} the existence of a bounded linear potential operator here is a non-trivial fact even in the Hilbert space case $p = 2$.

@@

Domains that admit finite triangulations allow for techniques \cye{using some form of passing through the triangulation and construction of the potentials step-by-step}. For scalar-valued functions in the context of Section~\ref{section:intro_grad}, Veeser and Verf\"urth~\cite{veeser2012poincare} provide computable upper bounds in the case of the classical Sobolev space $W^{1,p}(\Omega)$ over triangulated domains, with focus on efficient estimates for \cye{vertex} stars. 
%Naturally, these estimates deteriorates when the shape regularity of the mesh deteriorates. 
Such approaches were also previously used in the context of finite volume methods~\cite{Eym_Gal_Her_00}, broken (weakly continuous) Sobolev spaces~\cite{vohralik2005discrete}, or more recently in continuous--discrete comparison results~\cite{Brae_Pill_Sch_p_rob_09, ern2020stable, Chaum_Voh_p_rob_3D_H_curl_24,  Voh_loc_glob_H1_24}. Triangle by triangle, the potential is constructed over increasing subdomains, matching along the interfacing intersections, until the entire domain is exhausted.

In the gradient case of Section~\ref{section:intro_grad}, it is possible to perform a simple pass through the triangulation, where merely each subsequent simplex has to be connected to the previous ones by a $(n-1)$-dimensional face\cye{; face-connected triangulations are sufficient here}. We always construct a gradient potential over the new simplex, and since the gradient kernel is trivial, formed merely by constants, it is enough to fix this constant using the value already known on the $(n-1)$-dimensional face.

@@

\cye{We treat specifically the gradient operator (the setting of Section~\ref{section:intro_grad}) on triangulated domains that can be non-convex and non-star-shaped but where the triangulations are face connected (though not necessarily shellable). Here, Theorems~\ref{theorem:poincarefriedrichsestimate:grad} and~\ref{theorem:poincarefriedrichsestimate:gradzwei} are our main results.} 
\mwl{MV: remark from after Theorems~\ref{theorem:poincarefriedrichsestimate:grad} and~\ref{theorem:poincarefriedrichsestimate:gradzwei} moved here.} 
The computable Poincar\'e--Friedrichs constants obtained here depend on only a few parameters of the \cye{given} triangulation: the length of any traversal from the root simplex, the ratios of the volumes of \cye{the} adjacent simplices, the Poincar\'e--Friedrichs constants on each face patch, and the Poincar\'e--Friedrichs constant of the initial simplex. \cye{Here,} the \cye{face patch} Poincar\'e--Friedrichs constants \cye{are} estimated in terms of shape \cye{regularity parameters of the triangulation}; if the face patches of the triangulation are convex, then better estimates are possible.

\cye{Our main result for the curl and divergence operators (Section~\ref{section:intro_curl_div}) and more generally in the exterior calculus setting (Section~\ref{section:intro_EC}) for triangulated domains are Theorems~\ref{theorem:poincarefriedrichsestimate:exterior} and~\ref{theorem:poincarefriedrichsestimate:exterior:contraction}.
They treat triangulated domains that can be non-convex and non-star-shaped but where the triangulation is shellable. As discussed above, though this is slightly less generic than merely face-connected triangulations, this namely covers patches of simplices sharing a given vertex, edge, face, or $n$-simplex (stars) in two and three space dimensions, since these local patches are shellable. The dependencies of the obtained computable Poincar\'e--Friedrichs constants are similar to those of the gradient operator in Theorems~\ref{theorem:poincarefriedrichsestimate:grad} and~\ref{theorem:poincarefriedrichsestimate:gradzwei}.}

@@

\subsection{Analytical constants in Poincar\'e--Friedrichs inequalities over bounded convex domains} \label{section: PX_convex}

@@

The next auxiliary result establishes Poincar\'e--Friedrichs constants over face patches within simplicial triangulations. 
We emphasize that face patches are not necessarily convex, but we can still extend the results on convex domains from Section~\ref{section: PX_convex}. The following result is reasonably sharp when the two simplices have similar volumes and diameters.

@@

The discussion for the Bogovski\u{\i} operator is analogous. 

@@

        \item 
        We define a mapping $ \Xi_{1} : T \rightarrow U = \overline{\patch_{\calT}(S) \setminus T}$ by setting 
        \begin{align*}
            \Xi_{1} := \Psi \circ \Theta.
        \end{align*}
        It preserves the interface from $T$ onto $U$. 
        Let $\sigma_1(x) \geq \dots \geq \sigma_n(x)$ be the singular values of its Jacobian at $x \in T$. 
        Combining the previous estimates for $\Psi$ and $\Theta$ with interlacing inequalities for the singular values of product matrices provides 
        \begin{align*}
            \sigma_1(x) = \matnorm{ \Jacobian \Xi_{1} }
            &\leq 
            \Cfive{n}{k}(\calT)
            :=
            (k+1) 
            \diameterratio(\calT)
            \aspectratio(\calT)
            B_{T}^{2}
            \Big( 
                1 
                + 
                (k+1) 
                \left( 1 + B_{T} \right)
                \aspectratio(\calT)
            \Big)            
            ,
            \\
            \sigma_2(x), \dots, \sigma_n(x) 
            &\leq 
            \Cfiveprime{n}{k}(\calT)
            :=
            (k+1)B_{T}
            \diameterratio(\calT)
            \aspectratio(T) 
			\cdot 
			B_T
            \\
            \sigma_n(x)^{-1} = \matnorm{ \Jacobian \Xi_{1}^{-1} }
            &\leq
            \Csechs{n}{k}(\calT)
            :=
            (k+1) 
            \diameterratio(\calT)
            \aspectratio(\calT)
            B_{T}
            \Big( 
                1 
                + 
                (k+1) 
                \left( 1 + B_{T} \right)
                \aspectratio(\calT)
            \Big)
            \\
            \sigma_1(x)^{-1}, \dots, \sigma_{n-1}(x)^{-1}
            &\leq 
            \Csechsprime{n}{k}(\calT)
            :=
            (k+1)
            \diameterratio(\calT)
            \aspectratio(\calT) 
            \cdot 
			B_T
            .
        \end{align*}
        We get the determinant estimates
        \begin{align*}
            |\det(\Xi_{1})|
            &\leq 
            \Cdetfive{n}{k}(\calT)
            :=
            (k+1)^{n}
            \diameterratio(\calT)^{n}
            B_{T}^{n}
            \aspectratio(\calT)^{n}
            ,
            \\
            |\det(\Xi_{1}^{-1})|
            &\leq 
            \Cdetsechs{n}{k}(\calT)
            :=
            (k+1)^{n}
            \diameterratio(\calT)^{n}
            \aspectratio(\calT)^{n}
            .
        \end{align*}
        This completes the desired estimates for $\Xi_{1}$.

@@

\begin{lemma}\label{lemma:mixedbconsimplex:exteriorderivative}
    Let $T$ be an $n$-simplex 
    and let $F_{0},\dots,F_{\ell}$ be $\ell+1$ different faces of $T$. 
    Suppose that $u \in W^{p}\Lambda^{k}(T)$ such that 
    $\trace_{T,F_i} u = 0$ for $0 \leq i \leq \ell$.
    Then there exists $w \in W^{p}\Lambda^{k}(T)$ such that 
    $\trace_{T,F_i} w = 0$ for $0 \leq i \leq \ell$
    and  
    \begin{align*}
        \cartan w = \cartan u,
        \quad 
        \| w \|_{L^{p}(\Omega)} 
        \leq 
        C_{\PF,\Gamma,\ell,k,p}(T)
        \| \cartan u \|_{L^{p}(\Omega)}
        .
    \end{align*}
    Here, $C_{\PF,\Gamma,\ell,k,p}(T) > 0$ is a constant such that 
    \begin{align*}
        C_{\PF,\Gamma,\ell,k,p}(T)
        \leq 
        2^{\frac n p}
        2^{n+1} n A_{\Bogov}(n,k+1) \frac{ \vol_{}(B_1(0)) }{ (\ell+1)/n! } 
        \algebraicshapemeasure(T)^{k-1}
        \Ceins{n}(T)^{k}
        \diam(T)
        % \matnorm{\Jacobian \varphi}^{k}
        .
    \end{align*}
\end{lemma}
\mwl{MV: There is indeed $\diam(T)^{k}$, with the power $k$? Nothing like this appears in Lemma~\ref{lemma:PFexteriorsimplex}. Moreover, so far, all the constants $C_{{\PF},T,p}$, $C_{{\PF},T_1 \cup T_2,p}$ in Lemma~\ref{lemma:poincarefriedrichsoverfacepatch}, $C_{\PF,T,F,p}$ in Lemma~\ref{lemma:mixedbconsimplex}, or $C_{{\PF},k,\Omega,p}$ in Lemma~\ref{lemma:PFexteriorsimplex} had the diameter $\diam(T)$ inside, only the constants in Section~\ref{section: PX_convex} have it outside. Finally, $\Omega$ in the indices is missing, whereas $l$ is superfluous. We should unify.}
\begin{proof}
    There exists an affine bijection $\varphi : \hat T \rightarrow T$ 
    that maps the convex closure of the $n$ unit vectors onto the face $F_0$.
    We can also assume that the face of $\hat T$
    orthogonal to the $i$-th coordinate axis is mapped 
    onto the face $F_i$. 
    In what follows, we let $\tilde U$ be the domain obtained from reflecting $\hat T$ along the coordinate axes $l+1$ through $n$. 
    We see that $\vol(\tilde U) = (l+1)/n!$

@@

We settle the following notation for our main result. 
Whenever $\calT$ is an $n$-dimensional triangulation, we use the abbreviation
\begin{align*}
    C_{\PF,\Gamma,k,p}(\calT) 
    :=
    \max\limits_{ \substack{ T \in \calT \\ \dim(T)=n \\ 0 \leq l \leq n } }
    C_{\PF,\Gamma,l,k,p}(T)
    .
\end{align*}
Suppose that an $n$-dimensional triangulation $\calT$ has a shelling $T_{0}, T_{1}, \dots, T_{M}$. 
\begin{align*}
    \mwl{X_{m} := ...}
\end{align*}
For each $1 \leq m \leq M$, the interface of the new simplex to the previous intermediate triangulation is 
\begin{align*}
    \Gamma_{m} := T_{m} \cap \left( T_{0} \cup \dots \cup T_{m-1} \right)
\end{align*}
According to Lemma~\ref{lemma:existenceofstar}, 
there exists a simplex $S_{m} \in \calT$ 
such that $T_{m}$ is the last $n$-simplex in the shelling that belongs to $T_{m} \in \patch_{\calT}(S_{m})$.
In particular, $\patch_{\calT}(S_{m})$ is already included in the union of the simplices up to $T_m$. 
We write 
\begin{align*}
    \Minako_{m} := \patch_{\calT}(S_{m}) \setminus T_{m}.
\end{align*}
for the complement of $T_{m}$ in the star $\patch_{\calT}(S_{m})$.

@@

    We begin with the following preparatory observations. 
    Consider any $n$-simplex $T$ with vertices $v_0, v_1, \dots, v_n$, where $v_0$ is the origin. 
    We let $A_{T} \in \bbR^{n \times n}$ be the matrix that maps the $i$-th unit vector to $v_i$.
    That matrix maps the reference simplex $\Delta^{n}$ onto $T$. 
        F_T(x) 
        \frac{ \| A_{T}^{-1} x \|_{1} }{ \|x\|_2 } x
    maps $T$ onto a ``simplex with a curved face'' inside the Euclidean unit ball. Its inverse is
        F_T^{-1}(x) 
        \frac{ \|x\|_2 }{ \| A_{T}^{-1} x \|_{1} } x
    We analyze its Lipschitz properties.
    In what follows, $y \in \bbR^{n}$. 
    On the one hand, 
    we compute the Jacobian at any $x \in \mathring T$:
        \Jacobian F_T(x) \cdot y
        \frac{ \|x\| y - x \|x\|^{-1} x^t y }{\|x\|^2} 
        \left( \vecone \cdot A_{T}^{-1} x \right)
        \frac{x}{\|x\|} \left( \vecone \cdot A_{T}^{-1} y \right)
        \\&= 
        \left( \Id - \|x\|^{-2} x \otimes x^t \right)
        \frac{ 1 }{ \|x\| } 
        \left( \vecone \cdot A_{T}^{-1} x \right)
        \frac{x}{\|x\|} \left( \vecone \cdot A_{T}^{-1} x \right)
        \\&= 
        \left( A_{T}^{-t} \vecone \cdot \hat x \right)
        \left( \Id - \hat x \otimes \hat x^t \right) y
        \left( A_{T}^{-t} \vecone \cdot y \right) \hat x 
    This is an orthogonal sum of two vectors. Its norm satisfies:
        \| \Jacobian F_T(x) \cdot y \|^{2}
        \left| A_{T}^{-t} \vecone \cdot \hat x \right|^{2}
        \| \left( \Id - \hat x \otimes \hat x^t \right) y \|^{2}
        \left| A_{T}^{-t} \vecone \cdot y \right|^{2} \| \hat x \|^{2}
    The Euclidean operator norm satisfies 
        \| \Jacobian F_T(x) \| 
        \sqrt 2 \left\| A_{T}^{-t} \vecone \right\|_{2}
        \sqrt{2n} \cdot \sigma_{\max}(A_{T}^{-1}) 
    On the other hand, we can rewrite the Jacobian as:
        \Jacobian F_T(x) \cdot y
        \left( A_{T}^{-t} \vecone \cdot \hat x \right)
        y
        \left( A_{T}^{-t} \vecone \cdot \hat x \right)
        \left( \hat x \otimes \hat x^t \right) y
        \left( A_{T}^{-t} \vecone \cdot y \right) \hat x 
        \\&= 
        \left( A_{T}^{-t} \vecone \cdot \hat x \right)
        y
        \hat x 
            \left( A_{T}^{-t} \vecone \cdot \hat x \right) \hat x^{t}
            \vecone^{t} A_{T}^{-1} 
        \right) y
        \\&= 
        \left( A_{T}^{-t} \vecone \cdot \hat x \right)
        y
        \hat x 
            \left( A_{T}^{-t} \vecone \cdot \hat x \right) \hat x
            A_{T}^{-t} \vecone
        \right)^{t} y
    Its inverse is found via the Woodbury formula: 
    % Its inverse is easily verified to be:
        \Jacobian F_T(x)^{-1}
        \left( A_{T}^{-t} \vecone \cdot \hat x \right)^{-1}
        \left( A_{T}^{-t} \vecone \cdot \hat x \right)^{-2}
        \hat x 
            \left( A_{T}^{-t} \vecone \cdot \hat x \right) \hat x
            A_{T}^{-t} \vecone
        \right)^{t} 
        \\&= 
        \left( A_{T}^{-t} \vecone \cdot \hat x \right)^{-1}
            \hat x 
                % \left( A_{T}^{-t} \vecone \cdot \hat x \right)^{-1}
                % \left( A_{T}^{-t} \vecone \cdot \hat x \right) 
                \hat x
                \left( A_{T}^{-t} \vecone \cdot \hat x \right)^{-1}
                A_{T}^{-t} \vecone
            \right)^{t}
        \\&= 
        \left( \vecone \cdot A_{T}^{-1} \hat x \right)^{-1}
            \hat x 
                % \left( A_{T}^{-t} \vecone \cdot \hat x \right)^{-1}
                % \left( A_{T}^{-t} \vecone \cdot \hat x \right) 
                \hat x
                \left( \vecone \cdot A_{T}^{-1} \hat x \right)^{-1}
                A_{T}^{-t} \vecone
            \right)^{t}
        \\&= 
        \left( \vecone \cdot A_{T}^{-1} \hat x \right)^{-1}
            \Id - \hat x \otimes \hat x^{t}
            \left( \vecone \cdot A_{T}^{-1} \hat x \right)^{-1}
            \hat x \otimes \vecone^{t} A_{T}^{-1}
    The inverse acts via multiplication with $\left( \vecone \cdot A_{T}^{-1} \hat x \right)^{-1}$ on the span of $\hat x$.
    We notice the upper bound 
        \left( \vecone \cdot A_{T}^{-1} \hat x \right)^{-1}
        \| A_{T}^{-1} \hat x \|^{-1}_{1}
        \\&\leq  
        \| A_{T}^{-1} \hat x \|^{-1}_{2}
        \sigma_{\min}( A_{T}^{-1} )^{-1}
        \sigma_{\max}( A_{T} )
    This holds because $A_{T}^{-1} \hat x$ is in the positive quadrant.
    %, and its scalar product with $\vecone$ is minimized when it is colinear with a coordinate axis.
    If $y \in \bbR^{n}$ is orthogonal to $\hat x$, then 
        \| \Jacobian F_T(x)^{-1} \cdot y \|^{2}
%             \left( \vecone \cdot A_{T}^{-1} \hat x \right)^{-2}
%             \left\| 
%                 y
%             \right\|^{2}
%             \left( \vecone \cdot A_{T}^{-1} \hat x \right)^{-2}
%             \left| 
%                 \left( \vecone \cdot A_{T}^{-1} \hat x \right)^{-1}
%                 \vecone^{t} A_{T}^{-1} y 
%             \right|^{2}
%             \\&\leq 
        \left( \vecone \cdot A_{T}^{-1} \hat x \right)^{-2}
        \left\| 
            y
        \right\|^{2}
        \frac{ 
        \left| 
            \left( \vecone \cdot A_{T}^{-1} \hat x \right)^{-1}
            \vecone^{t} A_{T}^{-1} y 
        \right|^{2}
        }{
        \left( \vecone \cdot A_{T}^{-1} \hat x \right)^{2}
        }
    The operator norm of that inverse is:
        \| \Jacobian F_T(x)^{-1} \|
        \sigma_{\max}( A_{T} )
        \left( 1 + n \frac{ \sigma_{\max}(A_{T})^{2} }{ \sigma_{\min}(A_{T})^{2} } \right)^{\frac 1 2}
    Since the domain and codomain of $F_{T}$ are convex, 
    this also implies bounds on the Lipschitz constants of $F_{T}$ and its inverse. 
\color{red}
    On the one hand, given $x, y \in T$, 
    where $\| y \|_{1} \leq \| x \|_{1}$ without loss of generality, we find 
        \left\| F_T(x) - F_T(y) \right\|_{2}
        \left\| 
            \frac{ \| A_{T}^{-1} x \|_{1} }{ \|x\|_2 } x - \frac{ \| A_{T}^{-1} y \|_{1} }{ \|y\|_2 } y
        \right\|_{2}
        \left\| 
            \frac{ \| A_{T}^{-1} x \|_{1} }{ \|x\|_2 } x - \frac{ \| A_{T}^{-1} x \|_{1} }{ \|x\|_2 } y
        \right\|_{2}
        \left\| 
            \frac{ \| A_{T}^{-1} x \|_{1} }{ \|x\|_2 } y - \frac{ \| A_{T}^{-1} y \|_{1} }{ \|y\|_2 } y
        \right\|_{2}
        \frac{ \| A_{T}^{-1} x \|_{1} }{ \|x\|_2 } 
        \left\| 
            x - y
        \right\|_{2}
        \left| \|A_{T}^{-1} x\|_1 - \|A_{T}^{-1} y\|_1 \right|
        2\sqrt{n} \sigma_{\max}(A_{T}^{-1}) 
        \left\| x - y \right\|_{2}
    On the one hand, given $x, y \in T$, 
    where $\| x \|_{2} \leq \| y \|_{2}$ without loss of generality, we find 
        \left\| F_T(x) - F_T(y) \right\|_{2}
        \left\| 
            \frac{ \| A_{T}^{-1} x \|_{1} }{ \|x\|_2 } x - \frac{ \| A_{T}^{-1} y \|_{1} }{ \|y\|_2 } y
        \right\|_{2}
        \left\| 
            \frac{ \| A_{T}^{-1} x \|_{1} }{ \|x\|_2 } x - \frac{ \| A_{T}^{-1} x \|_{1} }{ \|x\|_2 } y
        \right\|_{2}
        \left\| 
            \frac{ \| A_{T}^{-1} x \|_{1} }{ \|x\|_2 } y - \frac{ \| A_{T}^{-1} y \|_{1} }{ \|x\|_2 } y
        \right\|_{2}
        \left\| 
            \frac{ \| A_{T}^{-1} y \|_{1} }{ \|x\|_2 } y - \frac{ \| A_{T}^{-1} y \|_{1} }{ \|y\|_2 } y
        \right\|_{2}
        \frac{ \| A_{T}^{-1} x \|_{1} }{ \|x\|_2 } 
        \left\| 
            x - y
        \right\|_{2}
        \left| \|A_{T}^{-1} x\|_1 - \|A_{T}^{-1} y\|_1 \right|
        \frac{\|y\|}{\|x\|}
        \|A_{T}^{-1} y\|_1 \|y\|_{2}
        \left| \frac{1}{\|x\|} - \frac{1}{\|y\|} \right|
        \frac{ \| A_{T}^{-1} x \|_{1} }{ \|x\|_2 } 
        \left\| 
            x - y
        \right\|_{2}
        \left| \|A_{T}^{-1} x\|_1 - \|A_{T}^{-1} y\|_1 \right|
        \frac{\|y\|}{\|x\|}

>>>>

%\subsection{\cye{Potential for the gradient operator and the corresponding Poincar\'e--Friedrichs inequality}} \label{subsection:intro_grad}

@@

\cye{Let a vector field $\bff$ that is a distributional curl of an object from $\bfW^{p}(\curl,\Omega)$, $\bff \in \curl \bfW^{p}(\curl,\Omega)$, and scalar-valued field $f$ that is a distributional divergence of an object from $\bfW^{p}(\divergence,\Omega)$, $f \in \divergence \bfW^{p}(\divergence,\Omega)$, be given.} 
We are interested in constants for inequalities such as 
\begin{subequations}\label{math:pot_PF_curl_div} \begin{align}
    \min\limits_{ \substack{ \bfu \in \bfW^{p}(\curl,\Omega) \\ \curl \bfu = \bff } } 
    \| \bfu \|_{L^{p}(\Omega)}
    &\leq 
    C_{\curl,\Omega,p}
    \| \bff \|_{L^{p}(\Omega)} \label{math:pot_PF_curl}
    ,
    \\
    \min\limits_{ \substack{ \bfu \in \bfW^{p}(\divergence,\Omega) \\ \divergence \bfu = f } } 
    \| \bfu \|_{L^{p}(\Omega)}
    &\leq 
    C_{\divergence,\Omega,p}
    \| f \|_{L^{p}(\Omega)} \label{math:pot_PF_div}
    ,
\end{align}\end{subequations}
and how these constants can be estimated by the operator norms of {\em linear} potential operators 
\begin{align*}
    &\Phi_{\curl} : \curl \bfW^{p}(\curl,\Omega) \rightarrow \bfW^{p}(\curl,\Omega)
    ,
    \\
    &\Phi_{\divergence} : \divergence \bfW^{p}(\divergence,\Omega) \rightarrow \bfW^{p}(\divergence,\Omega)
    .
\end{align*}
The fundamental difference \cye{with respect to Section~\ref{subsection:intro_grad}} is that the curl and divergence have infinite-dimensional kernels. In Section~\ref{subsection:intro_grad}, the kernel of the gradient is the one-dimensional space of constant functions. It is thus trivially complemented for all $p$, with a canonical choice of projection. By contrast, in the Banach space case $p \neq 2$, it is not immediately evident that the kernels of the curl and divergence operators are complemented, and a canonical projection only exists here in the Hilbert setting $p=2$. \cye{Actually,} the existence of a bounded linear potential operator here is a non-trivial fact even in the Hilbert space case $p = 2$.

@@

Domains that admit finite triangulations allow for techniques \cye{using some form of passing through the triangulation and construction of the potentials step-by-step}. For scalar-valued functions in the context of Section~\ref{subsection:intro_grad}, Veeser and Verf\"urth~\cite{veeser2012poincare} provide computable upper bounds in the case of the classical Sobolev space $W^{1,p}(\Omega)$ over triangulated domains, with focus on efficient estimates for \cye{vertex} stars. 
%Naturally, these estimates deteriorates when the shape regularity of the mesh deteriorates. 
Such approaches were also previously used in the context of finite volume methods~\cite{Eym_Gal_Her_00}, broken (weakly continuous) Sobolev spaces~\cite{vohralik2005discrete}, or more recently in continuous--discrete comparison results~\cite{Brae_Pill_Sch_p_rob_09, ern2020stable, Chaum_Voh_p_rob_3D_H_curl_24,  Voh_loc_glob_H1_24}. Triangle by triangle, the potential is constructed over increasing subdomains, matching along the interfacing intersections, until the entire domain is exhausted.

In the gradient case of Section~\ref{subsection:intro_grad}, it is possible to perform a simple pass through the triangulation, where merely each subsequent simplex has to be connected to the previous ones by a $(n-1)$-dimensional face\cye{; face-connected triangulations are sufficient here}. We always construct a gradient potential over the new simplex, and since the gradient kernel is trivial, formed merely by constants, it is enough to fix this constant using the value already known on the $(n-1)$-dimensional face.

@@

\cye{We treat specifically the gradient operator (the setting of Section~\ref{subsection:intro_grad}) on triangulated domains that can be non-convex and non-star-shaped but where the triangulations are face connected (though not necessarily shellable). Here, Theorem~\ref{theorem:poincarefriedrichsestimate:grad} is our main results.} 
\mwl{MV: remark from after Theorem~\ref{theorem:poincarefriedrichsestimate:grad}.} 
The computable Poincar\'e--Friedrichs constants obtained here depend on only a few parameters of the \cye{given} triangulation: the length of any traversal from the root simplex, the ratios of the volumes of \cye{the} adjacent simplices, the Poincar\'e--Friedrichs constants on each face patch, and the Poincar\'e--Friedrichs constant of the initial simplex. \cye{Here,} the \cye{face patch} Poincar\'e--Friedrichs constants \cye{are} estimated in terms of shape \cye{regularity parameters of the triangulation}; if the face patches of the triangulation are convex, then better estimates are possible.

\cye{Our main result for the curl and divergence operators (Section~\ref{section:intro_curl_div}) and more generally in the exterior calculus setting (Section~\ref{section:intro_EC}) for triangulated domains are Theorems~\ref{theorem:poincarefriedrichsestimate:exterior} and~\ref{theorem:poincarefriedrichsestimate:exterior:contraction}.
They treat triangulated domains that can be non-convex and non-star-shaped but where the triangulation is shellable. As discussed above, though this is slightly less generic than merely face-connected triangulations, this namely covers patches of simplices sharing a given vertex, edge, face, or $n$-simplex (stars) in two and three space dimensions, since these local patches are shellable. The dependencies of the obtained computable Poincar\'e--Friedrichs constants are similar to those of the gradient operator in Theorem~\ref{theorem:poincarefriedrichsestimate:grad}.}

@@

\subsection{Analytical constants in Poincar\'e--Friedrichs inequalities over bounded convex domains} \label{subsection: PX_convex}

@@

The next auxiliary result establishes Poincar\'e--Friedrichs constants over face patches within simplicial triangulations. 
We emphasize that face patches are not necessarily convex, but we can still extend the results on convex domains from Section~\ref{subsection: PX_convex}. The following result is reasonably sharp when the two simplices have similar volumes and diameters.

@@

Even though the discussion for the Bogovski\u{\i} operator is large analogous, some modifications are needed. 

@@

        \item 
        We define a mapping $ \Xi_{1} : T \rightarrow U = \overline{ \underlying{ \patch_{\calT}(S) } \setminus T}$ by setting 
        \begin{align*}
            \Xi_{1} := \Psi \circ \Theta.
        \end{align*}
        It preserves the interface from $T$ onto $U$. 
        Let $\sigma_1(x) \geq \dots \geq \sigma_n(x)$ be the singular values of its Jacobian at $x \in T$. 
        Combining the previous estimates for $\Psi$ and $\Theta$ with interlacing inequalities for the singular values of product matrices provides 
        \begin{align*}
            \sigma_1(x) = \matnorm{ \Jacobian \Xi_{1} }
            &\leq 
            \Cfive{n}{k}(\calT)
            :=
            (k+1) 
            \diameterratio(\calT)
            \aspectratio(\calT)
            B_{T}^{2}
            \Big( 
                1 
                + 
                (k+1) 
                \left( 1 + B_{T} \right)
                \aspectratio(\calT)
            \Big)            
            ,
            \\
            \sigma_2(x), \dots, \sigma_n(x) 
            &\leq 
            \Cfiveprime{n}{k}(\calT)
            :=
            (k+1)B_{T}
            \diameterratio(\calT)
            \aspectratio(T) 
			\cdot 
			B_T
            \\
            \sigma_n(x)^{-1} = \matnorm{ \Jacobian \Xi_{1}^{-1} }
            &\leq
            \Csechs{n}{k}(\calT)
            :=
            (k+1) 
            \diameterratio(\calT)
            \aspectratio(\calT)
            B_{T}
            \Big( 
                1 
                + 
                (k+1) 
                \left( 1 + B_{T} \right)
                \aspectratio(\calT)
            \Big)
            \\
            \sigma_1(x)^{-1}, \dots, \sigma_{n-1}(x)^{-1}
            &\leq 
            \Csechsprime{n}{k}(\calT)
            :=
            (k+1)
            \diameterratio(\calT)
            \aspectratio(\calT) 
            \cdot 
			B_T
            .
        \end{align*}
        We get the determinant estimates
        \begin{align*}
            |\det(\Xi_{1})|
            &\leq 
            \Cdetfive{n}{k}(\calT)
            :=
            (k+1)^{n}
            \diameterratio(\calT)^{n}
            B_{T}^{n}
            \aspectratio(\calT)^{n}
            ,
            \\
            |\det(\Xi_{1}^{-1})|
            &\leq 
            \Cdetsechs{n}{k}(\calT)
            :=
            (k+1)^{n}
            \diameterratio(\calT)^{n}
            \aspectratio(\calT)^{n}
            .
        \end{align*}
        This completes the desired estimates for $\Xi_{1}$.

@@

\begin{lemma}\label{lemma:mixedbconsimplex:exteriorderivative}
    Let $T$ be an $n$-simplex 
    and let $F_{0},\dots,F_{\ell}$ be $\ell+1$ different faces of $T$. 
    Suppose that $u \in W^{p}\Lambda^{k}(T)$ such that 
    $\trace_{T,F_i} u = 0$ for $0 \leq i \leq \ell$.
    Then there exists $w \in W^{p}\Lambda^{k}(T)$ such that 
    $\trace_{T,F_i} w = 0$ for $0 \leq i \leq \ell$
    and  
    \begin{align*}
        \cartan w = \cartan u,
        \quad 
        \| w \|_{L^{p}(\Omega)} 
        \leq 
        C_{\PF,\Gamma,\ell,k,p}(T)
        \| \cartan u \|_{L^{p}(\Omega)}
        .
    \end{align*}
    Here, $C_{\PF,\Gamma,\ell,k,p}(T) > 0$ is a constant such that 
    \begin{align*}
        C_{\PF,\Gamma,\ell,k,p}(T)
        \leq 
        2^{\frac n p}
        2^{n+1} n A_{\Bogov}(n,k+1) \frac{ \vol_{}(B_1(0)) }{ (\ell+1)/n! } 
        \algebraicshapemeasure(T)^{k-1}
        \Ceins{n}(T)^{k}
        \diam(T)
        % \matnorm{\Jacobian \varphi}^{k}
        .
    \end{align*}
\end{lemma}
\mwl{MV: There is indeed $\diam(T)^{k}$, with the power $k$? Nothing like this appears in Lemma~\ref{lemma:PFexteriorsimplex}. Moreover, so far, all the constants $C_{{\PF},T,p}$, $C_{{\PF},T_1 \cup T_2,p}$ in Lemma~\ref{lemma:poincarefriedrichsoverfacepatch}, $C_{\PF,T,F,p}$ in Lemma~\ref{lemma:mixedbconsimplex}, or $C_{{\PF},k,\Omega,p}$ in Lemma~\ref{lemma:PFexteriorsimplex} had the diameter $\diam(T)$ inside, only the constants in Section~\ref{subsection: PX_convex} have it outside. Finally, $\Omega$ in the indices is missing, whereas $l$ is superfluous. We should unify.}
\begin{proof}
    There exists an affine bijection $\varphi : \hat T \rightarrow T$ 
    that maps the convex closure of the $n$ unit vectors onto the face $F_0$.
    We can also assume that the face of $\hat T$
    orthogonal to the $i$-th coordinate axis is mapped 
    onto the face $F_i$. 
    In what follows, we let $\tilde U$ be the domain obtained from reflecting $\hat T$ along the coordinate axes $l+1$ through $n$. 
    We see that $\vol(\tilde U) = (l+1)/n!$

@@

We settle the following notation for our main result. 
Whenever $\calT$ is an $n$-dimensional triangulation, we use the abbreviation
\begin{align*}
    C_{\PF,\Gamma,k,p}(\calT) 
    :=
    \max\limits_{ \substack{ T \in \calT \\ \dim(T)=n \\ 0 \leq l \leq n } }
    C_{\PF,\Gamma,l,k,p}(T)
    .
\end{align*}
Suppose that an $n$-dimensional triangulation $\calT$ has a shelling $T_{0}, T_{1}, \dots, T_{M}$. 
For each $0 \leq m \leq M$, we write 
\begin{align*}
    X_{m} := T_0 \cup T_1 \cup \dots \cup T_{m}.
\end{align*}
For each $1 \leq m \leq M$, the interface of the new simplex to the previous intermediate triangulation is 
\begin{align*}
    \Gamma_{m} := T_{m} \cap X_{m-1} %\left( T_{0} \cup \dots \cup T_{m-1} \right)
	.
\end{align*}
According to Lemma~\ref{lemma:existenceofstar}, 
there exists a simplex $S_{m} \in \calT$ 
such that $T_{m}$ is the last $n$-simplex in the shelling that belongs to $T_{m} \in \patch_{\calT}(S_{m})$.
In particular, $\patch_{\calT}(S_{m})$ is already included in the union of the simplices up to $T_m$. 
We write 
\begin{align*}
    \Minako_{m} := \patch_{\calT}(S_{m}) \setminus T_{m}.
\end{align*}
for the complement of $T_{m}$ in the star $\patch_{\calT}(S_{m})$.

@@
